Refactoring the \texttt{nls()} function in R

### 2022-07-27 RefactoringNLS/RefactoringNLS.tex
Cleanup edits

The convergence criterion above, the principal termination control of
\texttt{nls()}, leaves out some possibilities that could be useful for
some problems. The package \texttt{nlsr}
((\protect\hyperlink{ref-nlsr-manual}{\textbf{nlsr-manual?}})) already
offers both the safeguarded relative offset test (\textbf{roffset}) as
well as a \textbf{small sum of squares} test (\textbf{smallsstest}) that
compares the latest evaluated sum of squared (weighted) residuals to
\texttt{e4} times the initial sum of squares, where
\texttt{e4\ \textless{}-\ (100*.Machine\$double.eps)\^{}4} is
approximately 2.43e-55.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{reala }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{; realb }\OtherTok{\textless{}{-}} \DecValTok{5}\NormalTok{; realc }\OtherTok{\textless{}{-}} \FloatTok{0.5}\NormalTok{; realr }\OtherTok{\textless{}{-}} \FloatTok{0.7}\NormalTok{; realm }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# underlying parameters}
\NormalTok{x }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{11} \CommentTok{\# x values; 11 data points}
\NormalTok{y }\OtherTok{\textless{}{-}}\NormalTok{ reala }\SpecialCharTok{+}\NormalTok{ realb }\SpecialCharTok{*}\NormalTok{ realr}\SpecialCharTok{\^{}}\NormalTok{(x }\SpecialCharTok{{-}}\NormalTok{ realm) }\SpecialCharTok{+}\NormalTok{ realc }\SpecialCharTok{*}\NormalTok{ x }\CommentTok{\# linear + exponential function}
\NormalTok{testdat }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(x, y) }\CommentTok{\# save in a data frame}
\NormalTok{strt }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{b =} \DecValTok{5}\NormalTok{, }\AttributeTok{c =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{r =} \FloatTok{0.7}\NormalTok{, }\AttributeTok{m =} \DecValTok{1}\NormalTok{) }\CommentTok{\# give programs a good start}
\NormalTok{jform }\OtherTok{\textless{}{-}}\NormalTok{ y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ a }\SpecialCharTok{+}\NormalTok{ b }\SpecialCharTok{*}\NormalTok{ r}\SpecialCharTok{\^{}}\NormalTok{(x }\SpecialCharTok{{-}}\NormalTok{ m) }\SpecialCharTok{+}\NormalTok{ c }\SpecialCharTok{*}\NormalTok{ x }\CommentTok{\# Formula}
\FunctionTok{library}\NormalTok{(nlsr)}
\NormalTok{linexp2 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(jform, }\AttributeTok{data =}\NormalTok{ testdat, }\AttributeTok{start =}\NormalTok{ strt, }\AttributeTok{trace =}\NormalTok{ F))}
\NormalTok{linexp2 }\CommentTok{\# Note singular values of Jacobian in rightmost column}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## nlsr object: x 
## residual sumsquares =  0  on  11 observations
##     after  1    Jacobian and  1 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## a                     -3            NA         NA         NA           0       26.49  
## b                      5            NA         NA         NA           0       9.615  
## c                    0.5            NA         NA         NA           0       2.466  
## r                    0.7            NA         NA         NA           0      0.1098  
## m                      1            NA         NA         NA           0   1.398e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# We get an error with nls()}
\NormalTok{linexp }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(jform, }\AttributeTok{data =}\NormalTok{ testdat, }\AttributeTok{start =}\NormalTok{ strt, }\AttributeTok{trace =}\NormalTok{ F))}
\end{Highlighting}
\end{Shaded}

@@

The functions \texttt{resid()} (an alias for \texttt{residuals()}) and
\texttt{fitted()} and \texttt{lhs()} are UNWEIGHTED. But if we return
\texttt{ans} from \texttt{nls()} or \texttt{minpack.lm::nlsLM} or our
new \texttt{nlsj} (interim package), then \texttt{ans\$m\$resid()} is
WEIGHTED. This needs clear documentation so users can get the answers
intended.

SHORT EXAMPLE??

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weed }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{5.308}\NormalTok{, }\FloatTok{7.24}\NormalTok{, }\FloatTok{9.638}\NormalTok{, }\FloatTok{12.866}\NormalTok{, }\FloatTok{17.069}\NormalTok{, }\FloatTok{23.192}\NormalTok{, }\FloatTok{31.443}\NormalTok{,}
          \FloatTok{38.558}\NormalTok{, }\FloatTok{50.156}\NormalTok{, }\FloatTok{62.948}\NormalTok{, }\FloatTok{75.995}\NormalTok{, }\FloatTok{91.972}\NormalTok{)}
\NormalTok{tt }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{12}
\NormalTok{weeddf }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(tt, weed)}
\NormalTok{wts }\OtherTok{\textless{}{-}} \FloatTok{0.5}\SpecialCharTok{\^{}}\NormalTok{tt }\CommentTok{\# simple weights}
\NormalTok{frmlogis }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym}\SpecialCharTok{/}\NormalTok{(}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}}\NormalTok{ tt)}\SpecialCharTok{/}\NormalTok{scal))}
\NormalTok{Asym}\OtherTok{\textless{}{-}}\DecValTok{1}\NormalTok{; xmid}\OtherTok{\textless{}{-}}\DecValTok{1}\NormalTok{; scal}\OtherTok{\textless{}{-}}\DecValTok{1}
\NormalTok{nowt}\OtherTok{\textless{}{-}}\FunctionTok{nls}\NormalTok{(weed }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(tt, Asym, xmid, scal))}
\NormalTok{nowt}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nowt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{inpar}\OtherTok{\textless{}{-}}\FunctionTok{getInitial}\NormalTok{(weed }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(tt, Asym, xmid, scal), }\AttributeTok{data=}\NormalTok{weeddf)}
\FunctionTok{print}\NormalTok{(inpar)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     Asym     xmid     scal 
## 196.1862  12.4173   3.1891
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(nlsr2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Attaching package: 'nlsr2'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:nlsr':
## 
##     codeDeriv, findSubexprs, fnDeriv, isCALL, isMINUSONE, isONE,
##     isZERO, model2rjfun, model2ssgrfun, modelexpr, newDeriv,
##     newSimplification, nlfb, nlsDeriv, nlsSimplify, nlxb, resgr, resss,
##     rjfundoc, sysDerivs, sysSimplifications, wrapnlsr
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nowtx}\OtherTok{\textless{}{-}}\FunctionTok{nlxb}\NormalTok{(frmlogis, }\AttributeTok{start=}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{1}\NormalTok{, }\AttributeTok{xmid=}\DecValTok{1}\NormalTok{, }\AttributeTok{scal=}\DecValTok{1}\NormalTok{), }\AttributeTok{trace=}\ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## <<lamda: 1e-04  SS= 23527 ( NA ) at  Asym = 1  xmid = 1  scal = 1 f/j  1 / 0
## <<lamda: 4e-05  SS= 10887 ( 0.8264 ) at  Asym = 52.337  xmid = 48.234  scal = 180.33 f/j  2 / 1
## <<lamda: 0.016  SS= 8948.4 ( 0.41672 ) at  Asym = 80.893  xmid = 48.394  scal = 169.52 f/j  6 / 2
## <<lamda: 0.0064  SS= 8853.5 ( 0.068832 ) at  Asym = 89.895  xmid = 66.542  scal = 118.27 f/j  7 / 3
## <<lamda: 0.0256  SS= 8673.7 ( 0.080434 ) at  Asym = 103.26  xmid = 69.707  scal = 91.474 f/j  9 / 4
## <<lamda: 0.1024  SS= 8563.7 ( 0.059227 ) at  Asym = 110.79  xmid = 67.869  scal = 81.345 f/j  11 / 5
## <<lamda: 0.04096  SS= 8392.1 ( 0.12852 ) at  Asym = 129.64  xmid = 62.354  scal = 48.413 f/j  12 / 6
## <<lamda: 0.16384  SS= 7777.5 ( 0.17269 ) at  Asym = 139.58  xmid = 48.216  scal = 39.581 f/j  14 / 7
## <<lamda: 0.65536  SS= 7547.4 ( 0.11307 ) at  Asym = 142.31  xmid = 43.841  scal = 33.748 f/j  16 / 8
## <<lamda: 0.26214  SS= 6212.9 ( 0.22183 ) at  Asym = 149.97  xmid = 28.123  scal = 17.451 f/j  17 / 9
## <<lamda: 1.0486  SS= 2919.3 ( 0.33274 ) at  Asym = 152.81  xmid = 15.368  scal = 7.5437 f/j  19 / 10
## <<lamda: 4.1943  SS= 1904 ( 0.74162 ) at  Asym = 153.41  xmid = 9.8392  scal = 1.8081 f/j  21 / 11
## <<lamda: 1.6777  SS= 113.91 ( 0.73923 ) at  Asym = 153.46  xmid = 10.631  scal = 2.5863 f/j  22 / 12
## <<lamda: 0.67109  SS= 11.223 ( 0.6629 ) at  Asym = 153.64  xmid = 10.972  scal = 2.8504 f/j  23 / 13
## <<lamda: 0.26844  SS= 10.289 ( 0.19356 ) at  Asym = 154.03  xmid = 11.02  scal = 2.8798 f/j  24 / 14
## <<lamda: 0.10737  SS= 9.8336 ( 0.11332 ) at  Asym = 154.95  xmid = 11.055  scal = 2.889 f/j  25 / 15
## <<lamda: 0.04295  SS= 8.8717 ( 0.16811 ) at  Asym = 157.04  xmid = 11.133  scal = 2.9084 f/j  26 / 16
## <<lamda: 0.01718  SS= 7.2069 ( 0.21397 ) at  Asym = 161.34  xmid = 11.29  scal = 2.9469 f/j  27 / 17
## <<lamda: 0.0068719  SS= 5.1659 ( 0.30108 ) at  Asym = 168.61  xmid = 11.548  scal = 3.0079 f/j  28 / 18
## <<lamda: 0.0027488  SS= 3.5835 ( 0.33934 ) at  Asym = 178.1  xmid = 11.867  scal = 3.079 f/j  29 / 19
## <<lamda: 0.0010995  SS= 2.8237 ( 0.28331 ) at  Asym = 187.23  xmid = 12.155  scal = 3.1385 f/j  30 / 20
## <<lamda: 0.0004398  SS= 2.6123 ( 0.16163 ) at  Asym = 193.29  xmid = 12.334  scal = 3.1734 f/j  31 / 21
## <<lamda: 0.00017592  SS= 2.588 ( 0.056549 ) at  Asym = 195.68  xmid = 12.403  scal = 3.1863 f/j  32 / 22
## <<lamda: 7.0369e-05  SS= 2.5873 ( 0.010514 ) at  Asym = 196.14  xmid = 12.416  scal = 3.1889 f/j  33 / 23
## <<lamda: 2.8147e-05  SS= 2.5873 ( 0.00090069 ) at  Asym = 196.18  xmid = 12.417  scal = 3.1891 f/j  34 / 24
## <<lamda: 1.1259e-05  SS= 2.5873 ( 3.4237e-05 ) at  Asym = 196.19  xmid = 12.417  scal = 3.1891 f/j  35 / 25
## <<lamda: 4.5036e-06  SS= 2.5873 ( 6.5557e-07 ) at  Asym = 196.19  xmid = 12.417  scal = 3.1891 f/j  36 / 26
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nowtx}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## nlsr2  object: x 
## residual sumsquares =  2.5873  on  12 observations
##     after  27    Jacobian and  36 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## Asym             196.186         11.31      17.35  3.167e-08  -3.432e-10       44.93  
## xmid             12.4173        0.3346      37.11  3.716e-11  -5.273e-08        15.6  
## scal             3.18908        0.0698      45.69  5.768e-12   1.752e-07      0.0474
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# nowtSSx0\textless{}{-}nlxb(weed \textasciitilde{} SSlogis(tt, Asym=1, xmid=1, scal=1), start=c(Asym=1, xmid=1, scal=1), trace=TRUE)}
\CommentTok{\# nowtSsx0}
\CommentTok{\# nowtSSx\textless{}{-}nlxb(weed \textasciitilde{} SSlogis(tt, Asym=1, xmid=1, scal=1), start=c(Asym=100, xmid=2, scal=1), trace=TRUE, }
\CommentTok{\#               control=list(japprox="jand"))}
\CommentTok{\# nowtSSx}
\NormalTok{nowtxi}\OtherTok{\textless{}{-}}\FunctionTok{nlxb}\NormalTok{(frmlogis, }\AttributeTok{start=}\NormalTok{inpar, }\AttributeTok{trace=}\ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## <<lamda: 1e-04  SS= 2.5873 ( NA ) at  Asym = 196.19  xmid = 12.417  scal = 3.1891 f/j  1 / 0
## <<lamda: 4e-05  SS= 2.5873 ( 9.4859e-07 ) at  Asym = 196.19  xmid = 12.417  scal = 3.1891 f/j  2 / 1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nowtxi}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## nlsr2  object: x 
## residual sumsquares =  2.5873  on  12 observations
##     after  2    Jacobian and  2 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## Asym             196.186         11.31      17.35  3.167e-08  -1.769e-09       44.93  
## xmid             12.4173        0.3346      37.11  3.716e-11  -2.902e-08        15.6  
## scal             3.18908        0.0698      45.69  5.768e-12   3.896e-08      0.0474
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usewt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nmod }\OtherTok{\textless{}{-}} \FunctionTok{nlsModel}\NormalTok{(frmlogis, }\AttributeTok{data=}\NormalTok{weeddf, }\AttributeTok{start=}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{1}\NormalTok{, }\AttributeTok{xmid=}\DecValTok{1}\NormalTok{, }\AttributeTok{scal=}\DecValTok{1}\NormalTok{), }\AttributeTok{wts=}\NormalTok{wts)}
\NormalTok{usewtx}\OtherTok{\textless{}{-}}\FunctionTok{nlxb}\NormalTok{(frmlogis,  }\AttributeTok{start=}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{1}\NormalTok{, }\AttributeTok{xmid=}\DecValTok{1}\NormalTok{, }\AttributeTok{scal=}\DecValTok{1}\NormalTok{), }\AttributeTok{weights=}\NormalTok{wts, }\AttributeTok{trace=}\ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## <<lamda: 1e-04  SS= 82.379 ( NA ) at  Asym = 1  xmid = 1  scal = 1 f/j  1 / 0
## <<lamda: 4e-05  SS= 27.961 ( 0.87591 ) at  Asym = 23.072  xmid = 24.605  scal = 33.424 f/j  2 / 1
## <<lamda: 0.0016  SS= 27.179 ( 0.10632 ) at  Asym = 27.088  xmid = 25.624  scal = 24.768 f/j  5 / 2
## <<lamda: 0.0064  SS= 26.326 ( 0.11129 ) at  Asym = 30.015  xmid = 24.489  scal = 21.971 f/j  7 / 3
## <<lamda: 0.00256  SS= 26.011 ( 0.18194 ) at  Asym = 36.922  xmid = 21.283  scal = 12.464 f/j  8 / 4
## <<lamda: 0.01024  SS= 21.446 ( 0.27496 ) at  Asym = 39.825  xmid = 16.241  scal = 10.675 f/j  10 / 5
## <<lamda: 0.04096  SS= 19.865 ( 0.18158 ) at  Asym = 40.684  xmid = 14.706  scal = 8.9148 f/j  12 / 6
## <<lamda: 0.016384  SS= 11.81 ( 0.32805 ) at  Asym = 42.981  xmid = 9.344  scal = 4.7286 f/j  13 / 7
## <<lamda: 0.065536  SS= 3.8548 ( 0.46972 ) at  Asym = 43.701  xmid = 6.3095  scal = 2.7564 f/j  15 / 8
## <<lamda: 0.026214  SS= 2.8386 ( 0.45514 ) at  Asym = 44.458  xmid = 5.503  scal = 2.0895 f/j  16 / 9
## <<lamda: 0.010486  SS= 2.4496 ( 0.2365 ) at  Asym = 46.866  xmid = 5.9479  scal = 2.3 f/j  17 / 10
## <<lamda: 0.0041943  SS= 2.0123 ( 0.21463 ) at  Asym = 51.378  xmid = 6.3516  scal = 2.3762 f/j  18 / 11
## <<lamda: 0.0016777  SS= 1.4462 ( 0.2703 ) at  Asym = 59.51  xmid = 7.03  scal = 2.5092 f/j  19 / 12
## <<lamda: 0.00067109  SS= 0.91272 ( 0.306 ) at  Asym = 71.787  xmid = 7.8952  scal = 2.6532 f/j  20 / 13
## <<lamda: 0.00026844  SS= 0.51816 ( 0.29993 ) at  Asym = 88.001  xmid = 8.8228  scal = 2.7857 f/j  21 / 14
## <<lamda: 0.00010737  SS= 0.26988 ( 0.25879 ) at  Asym = 107.49  xmid = 9.727  scal = 2.901 f/j  22 / 15
## <<lamda: 4.295e-05  SS= 0.13173 ( 0.20437 ) at  Asym = 129.22  xmid = 10.558  scal = 2.9978 f/j  23 / 16
## <<lamda: 1.718e-05  SS= 0.062773 ( 0.16799 ) at  Asym = 151.62  xmid = 11.279  scal = 3.0748 f/j  24 / 17
## <<lamda: 6.8719e-06  SS= 0.032464 ( 0.12324 ) at  Asym = 172.29  xmid = 11.849  scal = 3.1314 f/j  25 / 18
## <<lamda: 2.7488e-06  SS= 0.021923 ( 0.076781 ) at  Asym = 188.07  xmid = 12.235  scal = 3.1673 f/j  26 / 19
## <<lamda: 1.0995e-06  SS= 0.01977 ( 0.035429 ) at  Asym = 196.72  xmid = 12.43  scal = 3.1847 f/j  27 / 20
## <<lamda: 4.398e-07  SS= 0.019619 ( 0.0092109 ) at  Asym = 199.45  xmid = 12.49  scal = 3.19 f/j  28 / 21
## <<lamda: 1.7592e-07  SS= 0.019617 ( 0.00087595 ) at  Asym = 199.84  xmid = 12.499  scal = 3.1908 f/j  29 / 22
## <<lamda: 7.0369e-08  SS= 0.019617 ( 4.7993e-05 ) at  Asym = 199.86  xmid = 12.499  scal = 3.1909 f/j  30 / 23
## <<lamda: 2.8147e-08  SS= 0.019617 ( 8.2336e-07 ) at  Asym = 199.86  xmid = 12.499  scal = 3.1909 f/j  31 / 24
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usewtx}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## nlsr2  object: x 
## residual sumsquares =  0.019617  on  12 observations
##     after  25    Jacobian and  31 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## Asym             199.859         17.46      11.45  1.149e-06  -2.201e-11       7.858  
## xmid             12.4994        0.4131      30.26    2.3e-10    2.22e-09        0.79  
## scal             3.19088       0.04323      73.81  7.781e-14  -1.511e-08    0.002673
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# }\AlertTok{NOTE}\CommentTok{ jafwd and jaback fail!}
\NormalTok{usewtxn}\OtherTok{\textless{}{-}}\FunctionTok{nlxb}\NormalTok{(frmlogis,  }\AttributeTok{start=}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{1}\NormalTok{, }\AttributeTok{xmid=}\DecValTok{1}\NormalTok{, }\AttributeTok{scal=}\DecValTok{1}\NormalTok{), }\AttributeTok{weights=}\NormalTok{wts, }\AttributeTok{trace=}\ConstantTok{TRUE}\NormalTok{, }\AttributeTok{control=}\FunctionTok{list}\NormalTok{(}\AttributeTok{japprox=}\StringTok{"jacentral"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

>>>>

The convergence criterion above, the principal termination control of
\texttt{nls()}, leaves out some possibilities that could be useful for
some problems. The package \texttt{nlsr} (John C Nash and Duncan Murdoch
(\protect\hyperlink{ref-nlsr2019manual}{2019})) already offers both the
safeguarded relative offset test (\textbf{roffset}) as well as a
\textbf{small sum of squares} test (\textbf{smallsstest}) that compares
the latest evaluated sum of squared (weighted) residuals to \texttt{e4}
times the initial sum of squares, where
\texttt{e4\ \textless{}-\ (100*.Machine\$double.eps)\^{}4} is
approximately 2.43e-55. (NOTE: For the purposes of this article, we use
a developmental version of package \texttt{nlsr} labelled
\texttt{nlsr2}.)

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{reala }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{; realb }\OtherTok{\textless{}{-}} \DecValTok{5}\NormalTok{; realc }\OtherTok{\textless{}{-}} \FloatTok{0.5}\NormalTok{; realr }\OtherTok{\textless{}{-}} \FloatTok{0.7}\NormalTok{; realm }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# underlying parameters}
\NormalTok{x }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{11} \CommentTok{\# x values; 11 data points}
\NormalTok{y }\OtherTok{\textless{}{-}}\NormalTok{ reala }\SpecialCharTok{+}\NormalTok{ realb }\SpecialCharTok{*}\NormalTok{ realr}\SpecialCharTok{\^{}}\NormalTok{(x }\SpecialCharTok{{-}}\NormalTok{ realm) }\SpecialCharTok{+}\NormalTok{ realc }\SpecialCharTok{*}\NormalTok{ x }\CommentTok{\# linear + exponential function}
\NormalTok{testdat }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(x, y) }\CommentTok{\# save in a data frame}
\NormalTok{strt }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{b =} \DecValTok{5}\NormalTok{, }\AttributeTok{c =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{r =} \FloatTok{0.7}\NormalTok{, }\AttributeTok{m =} \DecValTok{1}\NormalTok{) }\CommentTok{\# give programs a good start}
\NormalTok{jform }\OtherTok{\textless{}{-}}\NormalTok{ y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ a }\SpecialCharTok{+}\NormalTok{ b }\SpecialCharTok{*}\NormalTok{ r}\SpecialCharTok{\^{}}\NormalTok{(x }\SpecialCharTok{{-}}\NormalTok{ m) }\SpecialCharTok{+}\NormalTok{ c }\SpecialCharTok{*}\NormalTok{ x }\CommentTok{\# Formula}
\FunctionTok{library}\NormalTok{(nlsr2)}
\NormalTok{linexp2 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(jform, }\AttributeTok{data =}\NormalTok{ testdat, }\AttributeTok{start =}\NormalTok{ strt, }\AttributeTok{trace =} \ConstantTok{FALSE}\NormalTok{))}
\NormalTok{linexp2 }\CommentTok{\# Note singular values of Jacobian in rightmost column}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## nlsr2  object: x 
## residual sumsquares =  0  on  11 observations
##     after  1    Jacobian and  1 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## a                     -3            NA         NA         NA           0       26.49  
## b                      5            NA         NA         NA           0       9.615  
## c                    0.5            NA         NA         NA           0       2.466  
## r                    0.7            NA         NA         NA           0      0.1098  
## m                      1            NA         NA         NA           0   1.398e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# We get an error with nls()}
\NormalTok{linexp }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(jform, }\AttributeTok{data =}\NormalTok{ testdat, }\AttributeTok{start =}\NormalTok{ strt, }\AttributeTok{trace =} \ConstantTok{FALSE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

The function \texttt{resid()} (an alias for \texttt{residuals()})
\texttt{nlsModel()} is WEIGHTED, as are those in the structured object
\texttt{m} returned by \texttt{nls()} or \texttt{minpack.lm::nlsLM},
e.g., \texttt{ans\$m\$resid()}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weed }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{5.308}\NormalTok{, }\FloatTok{7.24}\NormalTok{, }\FloatTok{9.638}\NormalTok{, }\FloatTok{12.866}\NormalTok{, }\FloatTok{17.069}\NormalTok{, }\FloatTok{23.192}\NormalTok{, }\FloatTok{31.443}\NormalTok{,}
          \FloatTok{38.558}\NormalTok{, }\FloatTok{50.156}\NormalTok{, }\FloatTok{62.948}\NormalTok{, }\FloatTok{75.995}\NormalTok{, }\FloatTok{91.972}\NormalTok{)}
\NormalTok{tt }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{12}
\NormalTok{weeddf }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(tt, weed)}
\NormalTok{wts }\OtherTok{\textless{}{-}} \FloatTok{0.5}\SpecialCharTok{\^{}}\NormalTok{tt }\CommentTok{\# simple weights}
\NormalTok{frmlogis }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym}\SpecialCharTok{/}\NormalTok{(}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}}\NormalTok{ tt)}\SpecialCharTok{/}\NormalTok{scal))}
\NormalTok{Asym}\OtherTok{\textless{}{-}}\DecValTok{1}\NormalTok{; xmid}\OtherTok{\textless{}{-}}\DecValTok{1}\NormalTok{; scal}\OtherTok{\textless{}{-}}\DecValTok{1}
\NormalTok{nowt}\OtherTok{\textless{}{-}}\FunctionTok{nls}\NormalTok{(weed }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(tt, Asym, xmid, scal)) }\CommentTok{\# UNWEIGHTED}
\NormalTok{nowt}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nowt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.}
\end{Highlighting}
\end{Shaded}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usewt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# WEIGHTED. Does NOT take coefficients.}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nmod0 }\OtherTok{\textless{}{-}} \FunctionTok{nlsModel}\NormalTok{(frmlogis, }\AttributeTok{data=}\NormalTok{weeddf, }\AttributeTok{start=}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{1}\NormalTok{, }\AttributeTok{xmid=}\DecValTok{1}\NormalTok{, }\AttributeTok{scal=}\DecValTok{1}\NormalTok{), }\AttributeTok{wts=}\NormalTok{wts)}
\NormalTok{nmod0}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# Parameters are supplied in nlsModel() \textasciigrave{}start\textasciigrave{} above.}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] 3.3998 3.2545 3.0961 2.9784 2.8438 2.7748 2.6910 2.3474 2.1724 1.9359
## [11] 1.6572 1.4214
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nmod }\OtherTok{\textless{}{-}} \FunctionTok{nlsModel}\NormalTok{(frmlogis, }\AttributeTok{data=}\NormalTok{weeddf, }\AttributeTok{start=}\FunctionTok{coef}\NormalTok{(usewt), }\AttributeTok{wts=}\NormalTok{wts)}
\NormalTok{nmod}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
##  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}
